---
title: "Describe and analyse factors in productivity"
output: md_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(GGally)
library(brms)
library(rstanarm)
library(bayesplot)
library(purrr)
library(extrafont)
library(ggsci)
library(ggpubr)
loadfonts()

# Set working directory to location of current file.
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Suppress dplyr summarise warning.
options(dplyr.summarise.inform = FALSE)
theme_set(theme_bw(base_size = 10))
theme_update(text = element_text(family = "Linux Libertine Display G"))

# Read in data, and drop the two rows that have no semrel value (-end and -iteur)
dat <- read.csv('../variables/sfx_data.csv') %>%
  drop_na()

# Standardise predictors to z-scores
dat <- dat %>% 
  mutate(z_jp = (mean_junc_prob - mean(mean_junc_prob)) / sd(mean_junc_prob),
         z_fr = (mean_log_freq_ratio - mean(mean_log_freq_ratio)) / sd(mean_log_freq_ratio),
         z_sm = (semrel_prob - mean(semrel_prob)) / sd(semrel_prob))
```


# Visualise variables

Plots of the variables on their original scales.

Dotplots with labels overlaid based on code here: https://stackoverflow.com/questions/44991607/how-do-i-label-the-dots-of-a-geom-dotplot-in-ggplot2/45001849#45001849

## Frequency ratio

```{r}
dotwidth <- 1

p_fr <- dat %>% 
  ggplot(aes(x=mean_log_freq_ratio)) +
  geom_dotplot(method='histodot',
               binwidth = dotwidth) +
   NULL

# Get y-positions of points plotted by geom_dotplot
# Warning: these positions are not given
point.pos <- ggplot_build(p_fr)$data[[1]]

# Order rows of mtcars by hp
idx <- order(dat$mean_log_freq_ratio)
dat2 <- dat[idx,]

# scale.fact needs fine tuning 
# It is strictly connected to the dimensions of the plot
scale.fact <- 0.1375
dat2$ytext <- point.pos$stackpos*scale.fact
dat2$xtext <- point.pos$x
lbls <- gsub(" ","\n",dat2$sfx)

ggplot(dat2, aes(mean_log_freq_ratio)) +
  geom_dotplot(method='histodot',
               fill="lightgrey",
               binwidth = dotwidth) +
  geom_text(aes(label=lbls, x=xtext, y=ytext), 
            family = "Linux Libertine Display G",
            size = 2.5) +
  theme(axis.ticks.y = element_blank(), 
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(x = 'Mean log frequency ratio',
       y = element_blank())

ggsave('imgs/fr.pdf', device=cairo_pdf, width=10, height=8, units='cm')
```

## Semantic relatedness

```{r}
dotwidth <- 0.03

p <- dat %>% 
  ggplot(aes(x=semrel_prob)) +
  geom_dotplot(method='histodot',
               binwidth = dotwidth) +
   NULL

# Get y-positions of points plotted by geom_dotplot
# Warning: these positions are not given
point.pos <- ggplot_build(p)$data[[1]]

# Order rows of mtcars by hp
idx <- order(dat$semrel_prob)
dat2 <- dat[idx,]

# scale.fact needs fine tuning 
# It is strictly connected to the dimensions of the plot
scale.fact <- 0.1375
dat2$ytext <- point.pos$stackpos*scale.fact
dat2$xtext <- point.pos$x
lbls <- gsub(" ","\n",dat2$sfx)

ggplot(dat2, aes(semrel_prob)) +
  geom_dotplot(method='histodot',
               fill="lightgrey",
               binwidth = dotwidth) +
  geom_text(aes(label=lbls, x=xtext, y=ytext), 
            family = "Linux Libertine Display G",
            size = 2.5) +
  theme(axis.ticks.y = element_blank(), 
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(x = 'Mean probability of semantic relatedness',
       y = element_blank())

ggsave('imgs/sm.pdf', device=cairo_pdf, width=10, height=8, units='cm')
```

## Junctural phonotactics

```{r}
dotwidth <- 0.004

p <- dat %>% 
  ggplot(aes(x=mean_junc_prob)) +
  geom_dotplot(method='histodot',
               binwidth = dotwidth) +
   NULL

# Get y-positions of points plotted by geom_dotplot
# Warning: these positions are not given
point.pos <- ggplot_build(p)$data[[1]]

# Order rows of mtcars by hp
idx <- order(dat$mean_junc_prob)
dat2 <- dat[idx,]

# scale.fact needs fine tuning 
# It is strictly connected to the dimensions of the plot
scale.fact <- 0.066
dat2$ytext <- point.pos$stackpos*scale.fact
dat2$xtext <- point.pos$x
lbls <- gsub(" ","\n",dat2$sfx)

ggplot(dat2, aes(mean_junc_prob)) +
  geom_dotplot(method='histodot',
               fill="lightgrey",
               binwidth = dotwidth) +
  geom_text(aes(label=lbls, x=xtext, y=ytext), 
            family = "Linux Libertine Display G",
            size = 2.3) +
  theme(axis.ticks.y = element_blank(), 
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(x = 'Mean probability of juncture appearing in simplexes',
       y = element_blank())

ggsave('imgs/jp.pdf', device=cairo_pdf, width=10, height=13, units='cm')
```


## Pairs plot of z-scores

```{r message = FALSE, warning = FALSE}
dat %>% 
  select(z_fr, z_sm, z_jp) %>% 
  rename('Frequency ratio\n(z-score)' = 'z_fr',
         'Sem. relatedness\n(z-score)' = 'z_sm',
         'Junc. phonotactics\n(z-score)' = 'z_jp') %>% 
  ggpairs(
    upper = list(combo = wrap('box_no_facet'),  # removes the box around the Corr cells in upper triangle
                 continuous = wrap('cor', family='Linux Libertine Display G'))  # no clue how, but can access font this way
  ) +
  theme(strip.text.x = element_text(size=10),
        strip.text.y = element_text(size=10))

ggsave('imgs/pairs.pdf', device=cairo_pdf, width=12, height=12, units='cm')
```

# Lognormal entropy

How well does a lognormal distribution approximate the distribution we observe for entropy?

If $y$ is log-normally distributed, then that means that $\log(y)$ is normally distributed.

```{r}
log_ent_mu <- mean(log(dat$entropy))
log_ent_sigma <- sd(log(dat$entropy))

dat %>% 
  ggplot(aes(x=log(entropy))) +
  geom_density(fill = 'red', colour='red', alpha=0.1) +
  stat_function(fun = dnorm, 
                n=101, 
                args = (list(mean=log_ent_mu, sd = log_ent_sigma)),
                colour = 'blue', size=1) +
  annotate(geom='text', x = 0.5, y = 0.6, label='Lognormal\napproximation', colour='blue') +
  annotate(geom='text', x = 1.5, y = 0.2, label='Entropy of\ntype frequency\ndistributions', colour='red') +
  NULL
```

# Prior prediction

```{r, message=FALSE, warning=FALSE, eval=FALSE}
fit_prior <- brm(entropy ~ z_jp + z_fr + z_sm,
  data = dat,
  family = lognormal(),
  prior = c(
    prior(normal(1, 0.5), class = Intercept),
    prior(normal(0, 0.3), class = sigma),
    prior(normal(0, 0.5), class = b)
  ),
  sample_prior = "only",
  iter = 3000
)

save(file='model_prior_pred.rda', fit_prior)
```


```{r, message=FALSE, warning=FALSE, eval=FALSE}
load('model_prior_pred.rda')
color_scheme_set('gray')
pp_check(fit_prior, type='boxplot', notch=FALSE, nsamples = 10, seed=1) +
  labs(y='Shannon entropy (bits)')
ggsave('imgs/ppcheck.pdf', device=cairo_pdf, width=12, height=6, units='cm')
```

# Sensitivity analysis

```{r eval=FALSE}
sds <- c(0.01, 0.05, 0.1, 0.5, 1)

sens_ana <- map_dfr(sds, function(sd){
  priorb <- paste0('normal(0, ', sd, ')')
  fit <- brm(entropy ~ z_jp + z_fr + z_sm,
             data = dat,
             family = lognormal(),
             prior = c(
               prior(normal(1, 0.5), class = Intercept),
               prior(normal(0, 0.3), class = sigma),
               prior_string(priorb, class = 'b')
             ),
             iter = 3000)
  posterior_summary(fit, pars = c('b_z_jp', 'b_z_fr', 'b_z_sm')) %>% 
    as_tibble(rownames=NA) %>% 
    rownames_to_column(var='param') %>% 
    mutate(prior = priorb)
})

write.csv(sens_ana, 'sens_ana.csv', row.names=FALSE)
```

```{r}
sens_ana <- read.csv('sens_ana.csv')

# Spell out and reorder variable names
levels(sens_ana$param) <- c('Frequency ratio', 'Junc. phonotactics', 'Sem. relatedness')
sens_ana$param <- factor(sens_ana$param, levels=c('Frequency ratio', 'Sem. relatedness', 'Junc. phonotactics'))

sens_ana$prior <- factor(sens_ana$prior, levels=rev(levels(sens_ana$prior)))  # Reverse order of factors

sens_ana %>% 
  ggplot(aes(x=prior, y=Estimate)) +
  geom_point() +
  geom_linerange(aes(ymin=Q2.5, ymax=Q97.5)) +
  facet_wrap(~param) +
  coord_flip() +
  geom_hline(yintercept=0, linetype='dashed') +
  labs(x = element_blank(),
       y = 'Posterior estimates for beta') +
  theme(strip.text.x = element_text(size=10)) +
  NULL

ggsave('imgs/sens.pdf', device=cairo_pdf, width=14, height=8, units='cm')
```


# Estimation

```{r, message=FALSE, warning=FALSE, eval=FALSE}
fit <- brm(entropy ~ z_jp + z_fr + z_sm,
  data = dat,
  family = lognormal(),
  prior = c(
    prior(normal(1, 0.5), class = Intercept),
    prior(normal(0, 0.3), class = sigma),
    prior(normal(0, 0.5), class = b)
  ),
  iter = 3000
)

summary(fit)

# pp_check(fit, type='boxplot', notch=FALSE, nsamples = 20)
save(file = 'model_fit.rda', fit)
```


```{r, message=FALSE, warning=FALSE}
load('model_fit.rda')
```

```{r}
summary(fit)
```


```{r, message=FALSE, warning=FALSE}
color_scheme_set('gray')
mcmc_areas(fit, 
           pars = c('b_z_jp', 'b_z_fr', 'b_z_sm'),
           prob = .95,
           point_est='mean') +
  geom_vline(xintercept=0, linetype='dashed') +
  scale_y_discrete(labels = c('b_z_jp' = 'Junc. phon.',
                            'b_z_fr' = 'Freq. ratio',
                            'b_z_sm' = 'Sem. rel.'),
                   limits = c('b_z_jp', 'b_z_sm', 'b_z_fr')) +
  coord_cartesian(ylim = c(1.4, 3.4)) +   # hack to move axis ticks down
  NULL
ggsave('imgs/beta-pstr.pdf', device=cairo_pdf, width=12, height=8, units='cm')
```


# Transforming LM estimates back to original (SD) scale

The above estimates are on the log scale.

For an interpretation of each of these effects in SDs, let us take for each parameter the difference in entropy between that parameter at its mean and at 1 SD below the mean, holding the other parameters at their means of zero (accomplished by leaving them out of the eq).

```{r}
int_samples <- posterior_samples(fit)$b_Intercept
b_z_jp_samples <- posterior_samples(fit)$b_z_jp
b_z_fr_samples <- posterior_samples(fit)$b_z_fr
b_z_sm_samples <- posterior_samples(fit)$b_z_sm

# JP
eff_jp_samples <- exp(int_samples) - exp(int_samples - b_z_jp_samples)
eff_jp <- c(mean = mean(eff_jp_samples), quantile(eff_jp_samples, c(.025, .975)))

# FR
eff_fr_samples <- exp(int_samples) - exp(int_samples - b_z_fr_samples)
eff_fr <- c(mean = mean(eff_fr_samples), quantile(eff_fr_samples, c(.025, .975)))

# SM
eff_sm_samples <- exp(int_samples) - exp(int_samples - b_z_sm_samples)
eff_sm <- c(mean = mean(eff_sm_samples), quantile(eff_sm_samples, c(.025, .975)))

bind_rows(list('Junc. phonotactics' = eff_jp, 'Freq. ratio' = eff_fr, 'Sem. relatedness' = eff_sm), .id = 'Factor')
```

```{r}
bind_rows(list('Junc. phonotactics' = eff_jp, 'Freq. ratio' = eff_fr, 'Sem. relatedness' = eff_sm), .id = 'Factor') %>% 
  xtable::xtable()
```


This means that, for example, an increase of one SD in junctural phonotactics near the middle of the range of that predictor will result in a decrease in entropy of -0.33 bits on average.
For frequency ratio, an increase of one SD will lead to an increase of 0.45 bits of entropy on average.
For semantic relatedness, one SD increase will lead to an increase in entropy of 0.45 bits on average.
In all of these cases, though, zero is included within the 95% credible interval.

# Bayes factors: SA with empirical data

Caveat: This code could look way nicer if R didn't keep crashing.

All of the H1 models are the same; they all contain all three predictors.
The H0 models have each predictor left out once.
So we can fit all the H1 models up front and use them below (a) for computing the empirical Bayes factor and (b) for posterior simulation to check the behaviour of BFs.

## H1 models

```{r}
prior_sd <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1)
```

```{r eval=FALSE}
for(psd in prior_sd){
  beta_prior <- paste0('normal(0,',psd,')')
  
  # Fit H1 model (all three predictors)
  h1 <- brm(entropy ~ z_jp + z_fr + z_sm,
            data = dat,
            family = lognormal(),
            prior = c(
              prior(normal(1, 0.5), class = Intercept),
              prior(normal(0, 0.3), class = sigma),
              set_prior(beta_prior, class = 'b')
            ),
            warmup = 2000,
            iter = 10000,
            cores = 1,
            save_pars = save_pars(all = TRUE))
  save(h1, file = paste0('bf_fits/emp_h1/psd', psd, '.rda'))
}
```

Now we look at the three factors one by one.

## Freq ratio

Fit models using each prior SD and save to `.rda`.
(The for loop structure somehow helps R crash less.)

```{r eval=FALSE}

for(psd in prior_sd){
  beta_prior <- paste0('normal(0,',psd,')')
  
  # Fit H0 model (freq ratio left out)
  h0 <- brm(entropy ~ z_jp + z_sm,
            data = dat,
            family = lognormal(),
            prior = c(
              prior(normal(1, 0.5), class = Intercept),
              prior(normal(0, 0.3), class = sigma),
              set_prior(beta_prior, class = 'b')
            ),
            warmup = 2000,
            iter = 10000,
            cores = 1,
            save_pars = save_pars(all = TRUE))
  save(h0, file = paste0('bf_fits/fr_emp_h0/psd', psd, '.rda'))
}

# nb: since they have the same names, will have to read in each one separately, compute bf,
# and then read in the next one so first one doesn't get overwritten
```

Read in each `.rda` file and estimate the Bayes factor for those two models using bridge sampling.
(`cores = 1` everywhere in an attempt not to avoid crashes.)

```{r eval=FALSE}
fr_bfs_list <- list()

for(psd in prior_sd){
  load(file = paste0('bf_fits/emp_h1/psd', psd, '.rda'))
  load(file = paste0('bf_fits/fr_emp_h0/psd', psd, '.rda'))

  lml_h1 <- bridge_sampler(h1, silent = TRUE, cores = 1)
  lml_h0 <- bridge_sampler(h0, silent = TRUE, cores = 1)

  fr_bfs_list[as.character(psd)] <- bayes_factor(lml_h1, lml_h0, cores=1)$bf
  
  message(paste('Done', psd))
}

rm(h0, h1, lml_h1, lml_h0)

fr_bfs <- bind_rows(fr_bfs_list, .id=sd) %>% 
  pivot_longer(cols=everything(), names_to='SD', values_to='BF10') %>% 
  mutate(SD = as.numeric(gsub('X', '', SD)))

write.csv(fr_bfs, 'bf_fits/fr_bfs_empirical.csv', row.names=FALSE)
```

```{r}
fr_bfs <- read.csv('bf_fits/fr_bfs_empirical.csv')
fr_bfs %>% 
  ggplot(aes(x=SD, y=BF10)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept=1, linetype='dashed') +
  scale_x_continuous(trans='log10', breaks = prior_sd) +
  scale_y_continuous(trans='log10') +
  theme(panel.grid.minor.x = element_blank()) +
  NULL
```

## Sem rel

Fit models using each prior SD and save to `.rda`.

```{r eval=FALSE}
for(psd in prior_sd){
  beta_prior <- paste0('normal(0,',psd,')')
  
  # Fit H0 model (semrel left out)
  h0 <- brm(entropy ~ z_jp + z_fr,
            data = dat,
            family = lognormal(),
            prior = c(
              prior(normal(1, 0.5), class = Intercept),
              prior(normal(0, 0.3), class = sigma),
              set_prior(beta_prior, class = 'b')
            ),
            warmup = 2000,
            iter = 10000,
            cores = 1,
            save_pars = save_pars(all = TRUE))
  save(h0, file = paste0('bf_fits/sm_emp_h0/psd', psd, '.rda'))
}
```

Read in each `.rda` file and estimate the Bayes factor for those two models using bridge sampling.

```{r eval=FALSE}
sm_bfs_list <- list()

for(psd in prior_sd){
  load(file = paste0('bf_fits/emp_h1/psd', psd, '.rda'))
  load(file = paste0('bf_fits/sm_emp_h0/psd', psd, '.rda'))

  lml_h1 <- bridge_sampler(h1, silent = TRUE, cores = 1)
  lml_h0 <- bridge_sampler(h0, silent = TRUE, cores = 1)

  sm_bfs_list[as.character(psd)] <- bayes_factor(lml_h1, lml_h0, cores=1)$bf
  
  message(paste('Done', psd))
}

rm(h0, h1, lml_h1, lml_h0)

sm_bfs <- bind_rows(sm_bfs_list, .id=sd) %>% 
  pivot_longer(cols=everything(), names_to='SD', values_to='BF10') %>% 
  mutate(SD = as.numeric(gsub('X', '', SD)))

write.csv(sm_bfs, 'bf_fits/sm_bfs_empirical.csv', row.names=FALSE)
```

```{r}
sm_bfs <- read.csv('bf_fits/sm_bfs_empirical.csv')
sm_bfs %>% 
  ggplot(aes(x=SD, y=BF10)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept=1, linetype='dashed') +
  scale_x_continuous(trans='log10', breaks = prior_sd) +
  theme(panel.grid.minor.x = element_blank()) +
  NULL
```

## Junc prob

Fit models using each prior SD and save to `.rda`.

```{r eval=FALSE}
for(psd in prior_sd){
  beta_prior <- paste0('normal(0,',psd,')')
  
  # Fit H0 model (junc prob left out)
  h0 <- brm(entropy ~ z_sm + z_fr,
            data = dat,
            family = lognormal(),
            prior = c(
              prior(normal(1, 0.5), class = Intercept),
              prior(normal(0, 0.3), class = sigma),
              set_prior(beta_prior, class = 'b')
            ),
            warmup = 2000,
            iter = 10000,
            cores = 1,
            save_pars = save_pars(all = TRUE))
  save(h0, file = paste0('bf_fits/jp_emp_h0/psd', psd, '.rda'))
}
```

Read in each `.rda` file and estimate the Bayes factor for those two models using bridge sampling.

```{r eval=FALSE}
jp_bfs_list <- list()

for(psd in prior_sd){
  load(file = paste0('bf_fits/emp_h1/psd', psd, '.rda'))
  load(file = paste0('bf_fits/jp_emp_h0/psd', psd, '.rda'))

  lml_h1 <- bridge_sampler(h1, silent = TRUE, cores = 1)
  lml_h0 <- bridge_sampler(h0, silent = TRUE, cores = 1)

  jp_bfs_list[as.character(psd)] <- bayes_factor(lml_h1, lml_h0, cores=1)$bf
  
  message(paste('Done', psd))
}

rm(h0, h1, lml_h1, lml_h0)

jp_bfs <- bind_rows(jp_bfs_list, .id=sd) %>% 
  pivot_longer(cols=everything(), names_to='SD', values_to='BF10') %>% 
  mutate(SD = as.numeric(gsub('X', '', SD)))

write.csv(jp_bfs, 'bf_fits/jp_bfs_empirical.csv', row.names=FALSE)
```

```{r}
jp_bfs <- read.csv('bf_fits/jp_bfs_empirical.csv')
jp_bfs %>% 
  ggplot(aes(x=SD, y=BF10)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept=1, linetype='dashed') +
  scale_x_continuous(trans='log10', breaks = prior_sd) +
  theme(panel.grid.minor.x = element_blank()) +
  NULL
```

## All plots

```{r}
bfs <- bind_rows(
  list(`Freq. ratio` = fr_bfs,
       `Sem. rel.` = sm_bfs,
       `Junc. phon.` = jp_bfs),
  .id = 'Factor'
)

# Spell out and reorder variable names
bfs$Factor <- as.factor(bfs$Factor)
bfs$Factor <- factor(bfs$Factor, levels=c('Freq. ratio', 'Sem. rel.', 'Junc. phon.'))

bfs %>% 
  ggplot(aes(x=SD, y=BF10, colour=Factor, shape=Factor)) +
  geom_hline(yintercept=1, linetype='dashed', size=0.3) +
  geom_point() +
  geom_line() +
  annotate(geom='text', family = "Linux Libertine Display G",
           label='Evidence in favour of H1', 
           y = 1.81, 
           x = .1) +
  annotate(geom='text', family = "Linux Libertine Display G",
           label='Evidence in favour of H0', 
           y = 0.2, 
           x = .1) +
  scale_x_continuous(trans='log10', breaks = prior_sd) +
  scale_y_continuous(trans='log10', limits = c(0.15, 1.85)) +
  theme(panel.grid.minor.x = element_blank()) +
  labs(colour = element_blank(),
       shape = element_blank(),
       x = 'Standard deviation of beta prior') +
  scale_color_grey() +
  NULL

ggsave('imgs/bfs.pdf', device=cairo_pdf, width=12, height=8, units='cm')
```